---
title: "Lab4a"
output: html_document
---

## Importing John Hopkins University Data

```{r}
library(tidyverse)
```

```{r}
report_03_11_2020 <- read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/03-11-2020.csv")) %>%
rename(Country_Region = "Country/Region", Province_State = "Province/State")
```


```{r}
library(data.table)
```


## 1. What are the differences between 3/13/2020, the Friday before spring break 2020 and yesterday 6 months later on 9/13/2020? 

```{r}
report_09_13_2020 <- read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/09-13-2020.csv"))
```

```{r}
str(report_03_11_2020)
```

```{r}
str(report_09_13_2020)
```

There are more column included in the 9/13/2020 data set, such as:  FIPS, Admin2, Active, Combined_Key, Incidence_Rate and Case-Fatality_Ratio.


## 2. Make a data frame with the total counts for each country by tibble.

```{r}
report_March_country_total <- report_03_11_2020 %>%
  group_by(Country_Region) %>%
  summarise(Total_Confirmed = sum(Confirmed),
    Total_Deaths = sum(Deaths),
    Total_Recovered = sum(Recovered)
  )
```


```{r}
report_Sept_country_total <- report_09_13_2020 %>%
  group_by(Country_Region) %>%
  summarise(
    Total_Confirmed = sum(Confirmed),
    Total_Deaths = sum(Deaths),
    Total_Recovered = sum(Recovered)
  )
```

```{r}
data.table(report_March_country_total)
```
```{r}
data.table(report_Sept_country_total)
```

## 3. To get counts for confirmed cases per U.S state use filter() to select US and group_by and summarize as above.

```{r}
report_March_US_conf <- report_03_11_2020 %>%
  filter(Country_Region == "US") %>%
  select(Province_State, Confirmed)

report_Sept_US_conf <- report_09_13_2020 %>%
  filter(Country_Region == "US") %>%
  group_by(Province_State) %>%
  summarize(Total_Confirmed = sum(Confirmed))
```


```{r}
data.table(report_March_US_conf)
```

```{r}
data.table(report_Sept_US_conf)
```
## 4.Use arrange_by() and slice() to get just the top 10 countries for Deaths.

```{r}
report_March_top10 <- report_March_country_total %>%
  select(Country_Region, Total_Deaths) %>%
  arrange(desc(Total_Deaths)) %>%
  slice(1:10)
report_Sept_top10 <- report_Sept_country_total %>%
  select(Country_Region, Total_Deaths) %>%
  arrange(desc(Total_Deaths)) %>%
  slice(1:10)
```


```{r}
data.table(report_March_top10)
```

```{r}
data.table(report_Sept_top10)
```

## 5. Make plots using geom_point() of the (top) 10 countries with the confirmed cases and deaths (two separate graphs) on 3/12/2020 and 9/12/2020.


```{r}
March_death_plot <- ggplot(data = report_March_top10, 
                       mapping = aes(x = Country_Region, y = Total_Deaths))
March_death_plot + 
    geom_point() +
  ggtitle("Total Deaths per Country for March 2020")
```



```{r}
Sept_death_plot <- ggplot(data = report_Sept_top10, 
                       mapping = aes(x = Country_Region, y = Total_Deaths))
Sept_death_plot + 
    geom_point() +
  ggtitle("Total Deaths per Country for September 2020")
```




```{r}
report_March_top10_conf <- report_March_country_total %>%
  select(Country_Region, Total_Confirmed) %>%
  arrange(desc(Total_Confirmed)) %>%
  slice(1:10)
report_Sept_top10_conf <- report_Sept_country_total %>%
  select(Country_Region, Total_Confirmed) %>%
  arrange(desc(Total_Confirmed)) %>%
  slice(1:10)

March_conf_plot <- ggplot(data = report_March_top10_conf, 
                       mapping = aes(x = Country_Region, y = Total_Confirmed))
March_conf_plot + 
    geom_point() +
  ggtitle("Top 10 Total Confirmed Cases per Country for March 2020")
```


```{r}
Sept_conf_plot <- ggplot(data = report_Sept_top10_conf, 
                       mapping = aes(x = Country_Region, y = Total_Confirmed))
Sept_conf_plot + 
    geom_point() +
  ggtitle("Top 10 Total Confirmed Cases per Country for September 2020")
```

## 6. Make using geom_bar of the (top) 10 states with the most confirmed cases and deaths (two separate graphs) on 3/12/2020 and 9/12/2020.

```{r}
report_March_top_US_conf <- report_March_US_conf %>%
  select(Province_State, Confirmed) %>%
  arrange(desc(Confirmed)) %>%
  slice(1:10)

ggplot(report_March_top_US_conf, aes(x = Confirmed, y = Province_State)) +
geom_bar(stat = 'identity') +
  ggtitle("Top 10 Total Confirmed Cases per US State for March 2020")
```

```{r}
report_Sept_top_US_conf <- report_Sept_US_conf %>%
  select(Province_State, Total_Confirmed) %>%
  arrange(desc(Total_Confirmed)) %>%
  slice(1:10)

ggplot(report_Sept_top_US_conf, aes(x = Total_Confirmed, y = Province_State)) +
geom_bar(stat = 'identity') +
  ggtitle("Top 10 Total Confirmed Cases per US State for September 2020")
```











