---
title: "Lab4b"
output: html_document
---

## Lab 4 - ggplot2 Exercises

```{r}
library(tidyverse)
```


```{r}
download.file("http://r-bio.github.io/data/surveys_complete.csv",
              "surveys_complete.csv")

surveys_complete <- read_csv(file = "surveys_complete.csv")
```

## Challenge 1. Sample plots. 

```{r}
# Assign plot to a variable
surveys_plot <- ggplot(data = surveys_complete, mapping = aes(x = weight, y = hindfoot_length))
```

```{r}
# Plot
surveys_plot + 
    geom_point()
```

## Challenge 1. Using the geom_hex() function.

```{r}
install.packages("hexbin")

library("hexbin")
```


```{r}
surveys_plot +
 geom_hex()
```

## Challenge 2. Use what you learned to create a scatter plot of weight over species_id with the plot types showing in different colors


```{r}
surveys <- read_csv("Lab3/portal_data_joined.csv")
```

```{r}
surveys_na_filtered <- surveys %>%
  filter(!is.na(weight))
```


```{r}
ggplot(data = surveys_na_filtered, 
        mapping = aes(x = species_id, y = weight)) +
  geom_point(aes(color = plot_type))
```


## Challenge 3a. Replace the box plot with a violin plot.

```{r}
surveys_ggplot <- surveys_complete %>%
  filter(!is.na(weight))
```


```{r}
ggplot(data = surveys_ggplot, mapping = aes(x = species_id, y = weight)) +
    geom_violin(alpha = 0)
```


## Challenge 3b. Represent weight in log10 scale.

```{r}
ggplot(data = surveys_ggplot, mapping = aes(x = species_id, y = weight)) +
  geom_violin(alpha = 0) +
  scale_y_log10()
```


## Challenge 3c. Create a boxplot for hindfoot_length. Overlay the boxplot layer on a jitter layer to show actual measurements.

```{r}
surveys_ggplot_filterna <- surveys_complete %>%
  filter(!is.na(hindfoot_length))

ggplot(data = surveys_ggplot_filterna, mapping = aes(x = species_id, y = hindfoot_length)) +
  geom_jitter(alpha = 0.3, color = "tomato") +
   geom_boxplot(alpha = 0)
```


## Challenge 3d. Add color to the data points on your boxplot according to the plot from which the sample was taken (plot_id).

```{r}
ggplot(data = surveys_ggplot_filterna, mapping = aes(x = species_id, y = hindfoot_length)) +
  geom_jitter(alpha = 0.3, aes(color = plot_id)) +
   geom_boxplot(alpha = 0)
```

## Challenge 4. If plot_id was a factor.

```{r}
surveys_ggplot_filterna <- mutate_at(surveys_ggplot_filterna, vars(plot_id), as.factor)

ggplot(data = surveys_ggplot_filterna, mapping = aes(x = species_id, y = hindfoot_length)) +
  geom_jitter(alpha = 0.3, aes(color = plot_id)) +
   geom_boxplot(alpha = 0)
```












