---
title: "Lab4b"
output: html_document
---

## Lab 4 - ggplot2 Exercises

```{r}
library(tidyverse)
```


```{r}
download.file("http://r-bio.github.io/data/surveys_complete.csv",
              "surveys_complete.csv")

surveys_complete <- read_csv(file = "surveys_complete.csv")
```

#Example Plot


```{r}
# Assign plot to a variable
surveys_plot <- ggplot(data = surveys_complete, mapping = aes(x = weight, y = hindfoot_length))
```


```{r}
# Draw the plot
surveys_plot + 
    geom_point()
```

## Use the geom_hex() function.

```{r}
install.packages("hexbin")

library("hexbin")
```


```{r}
surveys_plot +
 geom_hex()
```

## Use what you learned to create a scatter plot of weight over species_id with the plot types showing in different colors

```{r}
surveys_ggplot <- surveys_complete %>%
  filter(!is.na(weight))

```


```{r}
ggplot(data = surveys_ggplot, 
        mapping = aes(x = species_id, y = weight)) +
  geom_point(alpha = 0.1)
```



## Replace the box plot with a violin plot

```{r}
ggplot(data = surveys_ggplot, mapping = aes(x = species_id, y = weight)) +
    geom_violin(alpha = 0)
```

## Represent weight in log10 scale

```{r}
ggplot(data = surveys_ggplot, mapping = aes(x = species_id, y = weight)) +
  geom_violin(alpha = 0) +
  scale_y_log10()
```


## Create a boxplot for hindfoot_length. Overlay the boxplot layer on a jitter layer to show actual measurements.

```{r}
surveys_ggplot_2 <- surveys_complete %>%
  filter(!is.na(hindfoot_length))

ggplot(data = surveys_ggplot_2, mapping = aes(x = species_id, y = hindfoot_length)) +
  geom_jitter(alpha = 0.3, color = "tomato") +
   geom_boxplot(alpha = 0)
```


## Add color to the data points on your boxplot according to the plot from which the sample was taken (plot_id).

```{r}
ggplot(data = surveys_ggplot_2, mapping = aes(x = species_id, y = hindfoot_length)) +
  geom_jitter(alpha = 0.3, aes(color = plot_id)) +
   geom_boxplot(alpha = 0)
```

## If plot_id was a factor

```{r}
#set as factor 
surveys_ggplot_2 <- mutate_at(surveys_ggplot_2, vars(plot_id), as.factor)
#plot
ggplot(data = surveys_ggplot_2, mapping = aes(x = species_id, y = hindfoot_length)) +
  geom_jitter(alpha = 0.3, aes(color = plot_id)) +
   geom_boxplot(alpha = 0)
```












