---
title: "September 26, 2020 COVID Report for the US and Oklahoma"
author: "Kaitlyn Gilliam "
output: html_document
bibliography: references_Lab6.bib
---

```{r, include = FALSE}
install.packages("tidyverse")
library(tidyverse)
library(lubridate)
install.packages("plotly")
library(plotly)
```


```{r, include = FALSE}
install.packages("maps")
install.packages("mapdata")
install.packages("viridis")
install.packages("wesanderson")
library(tidyverse)
library(maps)
library(mapdata)
library(lubridate)
library(viridis)
library(wesanderson)
```


```{r, include = FALSE}
# Read in the daily report
daily_report_drb <- read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/09-26-2020.csv")) %>%
  rename(Long = "Long_") %>% 
  group_by(Country_Region) %>% 
  summarize(Confirmed = sum(Confirmed), Deaths = sum(Deaths))

# Read in the world map data
world <- as_tibble(map_data("world"))

# Check to see if there are differences in the naming of countries
setdiff(world$region, daily_report_drb$Country_Region) 
```



## World Report of COVID-19 Deaths on 9-26-2020

```{r, message = FALSE, warning = FALSE, error = FALSE, echo = FALSE}
# Many of these countries are considered states or territories in the JHU covid reports
world <- as_tibble(map_data("world")) %>% 
 mutate(region = str_replace_all(region, c("USA" = "US", "Czech Republic" = "Czechia",  
        "Ivory Coast" = "Cote d'Ivoire", "Democratic Republic of the Congo" = "Congo (Kinshasa)", 
        "Republic of Congo" = "Congo (Brazzaville)")))

# Join the covid report with the map data
country_join <- left_join(world, daily_report_drb, by = c("region" = "Country_Region"))

# Create the graph
ggplotly(
ggplot(data = world, mapping = aes(x = long, y = lat, text = region, group = group)) + 
  coord_fixed(1.3) + 
# Add data layer
  geom_polygon(data = country_join, aes(fill = Deaths), color = "black") +
  scale_fill_gradientn(colours = 
                         wes_palette("Zissou1", 100, type = "continuous")) +
  labs(title = "COVID-19 Deaths 9-26-2020")
)

```


## World Report of COVID-19 Confirmed Cases on 9-26-2020 

```{r, include = FALSE}
daily_report_new <- read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/09-26-2020.csv")) %>% 
  rename(Long = "Long_") %>% 
  filter(Lat != "NA") %>% 
  filter(Long != "NA") 
```


```{r, message = FALSE, warning = FALSE, error = FALSE, echo = FALSE}
ggplot(daily_report_new, aes(x = Long, y = Lat, size = Confirmed/10000)) +
    borders("world", colour = NA, fill = "grey90") +
    theme_bw() +
    geom_point(shape = 21, color='blue', fill='blue', alpha = 0.5) +
    labs(title = 'COVID-19 Confirmed Cases 9-26-2020',x = 'lat', y = 'long',
        size="Cases (x10000))") +
    theme(legend.position = "right") +
    coord_fixed(ratio=1.5)
```

## United States Confirmed COVID-19 Cases on 9-26-2020

```{r, message = FALSE, warning = FALSE, error = FALSE, echo = FALSE}
daily_report_anisa <- read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/09-26-2020.csv")) %>% 
  rename(Long = "Long_") %>% 
  filter(Country_Region == "US") %>% 
  filter (!Province_State %in% c("Alaska","Hawaii", "American Samoa",
                  "Puerto Rico","Northern Mariana Islands", 
                  "Virgin Islands", "Recovered", "Guam", "Grand Princess",
                  "District of Columbia", "Diamond Princess")) %>% 
  filter(Lat > 0)
```

```{r, message = FALSE, warning = FALSE, error = FALSE, echo = FALSE}
mybreaks <- c(1, 100, 1000, 10000, 10000)
ggplot(daily_report_anisa, aes(x = Long, y = Lat, size = Confirmed)) +
    borders("state", colour = "white", fill = "grey90") +
    geom_point(aes(x=Long, y=Lat, size=Confirmed, color=Confirmed),stroke=F, alpha=0.7) +
    scale_size_continuous(name="Cases", trans="log", range=c(1,7), 
                        breaks=mybreaks, labels = c("1-99",
                        "100-999", "1,000-9,999", "10,000-99,999", "50,000+")) +
    scale_color_viridis_c(option="viridis",name="Cases",
                        trans="log", breaks=mybreaks, labels = c("1-99",
                        "100-999", "1,000-9,999", "10,000-99,999", "50,000+"))  +
# Cleaning up the graph
  
  theme_void() + 
    guides( colour = guide_legend()) +
    labs(title = "COVID-19 Confirmed Cases in the US 9-26-2020") +
    theme(
      legend.position = "bottom",
      text = element_text(color = "#22211d"),
      plot.background = element_rect(fill = "#ffffff", color = NA), 
      panel.background = element_rect(fill = "#ffffff", color = NA), 
      legend.background = element_rect(fill = "#ffffff", color = NA)
    ) +
    coord_fixed(ratio=1.5)
```


## Confirmed COVID-19 US Cases by County on 9-26-2020 


```{r, message = FALSE, warning = FALSE, error = FALSE, echo = FALSE}
library(RColorBrewer)
report_county <- read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/09-26-2020.csv")) %>% 
  rename(Long = "Long_") %>% 
  unite(Key, Admin2, Province_State, sep = ".") %>% 
  group_by(Key) %>% 
  summarize(Confirmed = sum(Confirmed)) %>% 
  mutate(Key = tolower(Key))
# get and format the map data
us <- map_data("state")
counties <- map_data("county") %>% 
  unite(Key, subregion, region, sep = ".", remove = FALSE)
# Join the 2 tibbles
state_join <- left_join(counties, report_county, by = c("Key"))
# sum(is.na(state_join$Confirmed))
ggplot(data = us, mapping = aes(x = long, y = lat, group = group)) + 
  coord_fixed(1.3) + 
  # Add data layer
    borders("state", colour = "black") +
  geom_polygon(data = state_join, aes(fill = Confirmed)) +
  scale_fill_gradientn(colors = brewer.pal(n = 5, name = "BuGn"),
                       breaks = c(1, 10, 100, 1000, 10000, 100000),
                       trans = "log10", na.value = "White") +
  ggtitle("Number of Confirmed Cases by US County 9-26-2020") +
  theme_bw() 
```

## COVID-19 Confirmed Cases by Oklahoma County on 9-26-2020

```{r, message = FALSE, warning = FALSE, error = FALSE, echo = FALSE}
report_oklahoma <- read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/09-26-2020.csv")) %>% 
  rename(Long = "Long_") %>% 
  filter(Province_State == "Oklahoma") %>% 
  group_by(Admin2) %>% 
  summarize(Confirmed = sum(Confirmed)) %>% 
  mutate(Admin2 = tolower(Admin2))
us <- map_data("state")
ok_us <- subset(us, region == "oklahoma")
counties <- map_data("county")
ok_county <- subset(counties, region == "oklahoma")
state_join <- left_join(ok_county, report_oklahoma, by = c("subregion" = "Admin2")) 
# plot state map
ggplot(data = ok_county, mapping = aes(x = long, y = lat, group = group)) + 
  coord_fixed(1.3) + 
# Add data layer
  geom_polygon(data = state_join, aes(fill = Confirmed), color = "white") +
    scale_fill_gradientn(colors = brewer.pal(n = 5, name = "PuRd"),
                         trans = "log10") +
  labs(title = "COVID-19 Confirmed Cases in Oklahoma County on 9-26-2020")


```

## Interactive graph displaying the confirmed cases by Oklahoma county on 9-26-2020

```{r, message = FALSE, warning = FALSE, error = FALSE, echo = FALSE}
library(plotly)
ggplotly(
  ggplot(data = ok_county, mapping = aes(x = long, y = lat, group = group)) + 
  coord_fixed(1.3) + 
# Add data layer
  geom_polygon(data = state_join, aes(fill = Confirmed), color = "black") +
  scale_fill_gradientn(colours = 
                         wes_palette("GrandBudapest2", 100, type = "continuous")) +

  ggtitle("COVID-19 Confirmed Cases in Oklahoma 9-26-2020") +
# Cleaning up the graph
  labs(x=NULL, y=NULL) +
  theme(panel.border = element_blank()) +
  theme(panel.background = element_blank()) +
  theme(axis.ticks = element_blank()) +
  theme(axis.text = element_blank())
)
```


Application written in R [@RCoreTeam] using the Shiny framework [@Chang2015].

Data obtained from the COVID-19 Data Repository by the Center for Systems Science and Engineering (CSSE) at Johns Hopkins University (Dong, et al, 2020). 

## REFERENCES
































