---
title: "Lab9a"
author: Kaitlyn Gilliam 
output: html_document
---

```{r}
library(stringr)
library(tidyverse)
```


## 14.3.1 Basic matches  
1. Explain why each of these strings don’t match a \: "\", "\\", "\\\". 
Answer: They are incomplete expressions.  

2. How would you match the sequence "'\ ?

```{r}
x <- "`\\"
writeLines(x)
```

```{r}
str_view(x, "'\\\\")
```


3. What patterns will the regular expression \..\..\.. match? How would you represent it as a string?

```{r}
b <- c("\\..\\..\\..")
writeLines(b)
```


```{r}
str_view(b, "\\..")
```


## 14.3.2 Anchors

1. How would you match the literal string "$^$"?

```{r}
a <- c("$^$", "$")
str_view(a, "\\$\\^\\$")
```

## 14.3.3 Character classes and alternatives

1. Create regular expressions to find all words that:

    1. Start with a vowel.
    
```{r}
c <- c("apple", "egg", "boat", "bed", "wise", "hmm", "psst")
str_view(c, "^[aeiouy]")
```
    
    2. That only contain consonants. (Hint: thinking about matching “not”-vowels.)

```{r}
str_view(c, "^[^aeiou]*$")
```

    3. End with ed, but not with eed.
   
```{r}
str_view(c, "[^e]ed$")
```

    4. End with ing or ise.

```{r}
str_view(c, "(ise|ing)$")
```


## Exercise 14.3.4 Repetition

1. Describe the equivalents of ?, +, * in {m,n} form.

```{r}
##  ?: 0 or 1
##  +: 1 or more
##  *: 0 or more
#### {0,1}
#### {1,}
#### {0,}
```

## Exercise 14.3.5 Grouping and backreferences

1. Describe, in words, what these expressions will match:
    1. (.)\1\1
    2. "(.)(.)\\2\\1"
    3. (..)\1
    4. "(.).\\1.\\1"
    5. "(.)(.)(.).*\\3\\2\\1"

```{r}
fruits <- c("banana", "coconut", "cucumber", "jujube", "papaya", "salal berry")

# 1. A character repeated three times.
  str_view(fruits, "(.)\\1\\1")
# 2. A pair of characters followed by the pair in reverse.
  str_view(fruits, "(.)(.)\\2\\1")
# 3. Any pair of characters repeated.
  str_view(fruits, "(..)\\1")
# 4. A character followed by any character, then the first character, any character, then the first character finally.
  str_view(fruits, "(.).\\1.\\1")
# 5. Three of any characters, followed by an unknown number of characters, but that followed by the first three characters in reverse order.
  str_view(fruits, "(.)(.)(.).*\\3\\2\\1")
```

## 14.4.1 Detect matches

1. Find all words that start or end with x.

```{r}
words[str_detect(words, "^x|x$")]
start_x <- str_detect(words, "^x") 
end_x <- str_detect(words, "x$")
words[start_x | end_x]
```

2. Words starting with vowel and ending with consonant.

```{r}
words[str_detect(words, "^[aeiou].*[^aeiou]$")]
start_vowel <- str_detect(words, "^[aeiou]") 
end_consonant <- str_detect(words , "[^aeiou]$")
words[start_vowel & end_consonant]
```

3. Are there any words that contain at least one of each different vowel?

```{r}
words[str_detect(words, "a") &
  str_detect(words, "e") &
  str_detect(words, "i") &
  str_detect(words, "o") &
  str_detect(words, "u")]
```

## 14.4.2 Extract matches 

1. In the previous example, you might have noticed that the regular expression matched “flickered”, which is not a colour. Modify the regex to fix the problem.

```{r}
colours <- c("red", "orange", "yellow", "green", "blue", "purple")
colour_words <- str_c("\\b(", str_c(colours, collapse = "|"), ")\\b")
more <- sentences[str_count(sentences, colour_words) > 1]
str_view_all(more, colour_words)
```

## 14.4.3 Grouped matches 

1. Find all words that come after a “number” like “one”, “two”, “three” etc. Pull out both the number and the word.

```{r}
numword <- "\\b(one|two|three|four|five|six|seven|eight|nine|ten) +(\\w+)"
sentences[str_detect(sentences, numword)] %>%
  str_extract(numword)
```

## 14.4.4 Replacing matches 

1. Replace all forward slashes in a string with backslashes.

```{r}
pronouns <- str_replace_all("he/him/his", "/", "\\\\")
writeLines(pronouns)
```

## Exercise 14.4.5 Splitting

1. Split up a string like “apples, pears, and bananas” into individual components.

```{r}
fruit <- c("apples, pears, and bananas") 
str_split(fruit, boundary("word"))[[1]]
```






