---
title: "Lab3b"
output: html_document
---

```{r}
library("tidyverse")
```


```{r}
surveys <- read_csv("portal_data_joined.csv")
```

```{r}
str(surveys)
```

## Challenge_1

Using pipes, subset the surveys data to include animals collected before 1995 and retain only the columns year, sex, and weight.

```{r}
surveys %>%
    filter(year < 1995) %>%
    select(year, sex, weight)
```
## Challenge_2

Create a new data frame from the surveys data that meets the following criteria: contains only the species_id column and a new column called hindfoot_cm containing the hindfoot_length values converted to centimeters. In this hindfoot_cm column, there are no NAs and all values are less than 3.

```{r}
surveys_hindfoot_cm <- surveys %>%
    filter(!is.na(hindfoot_length)) %>%
    mutate(hindfoot_cm = hindfoot_length / 10) %>%
    filter(hindfoot_cm < 3) %>%
    select(species_id, hindfoot_cm)
```

## Challenge_3

1. How many animals were caught in each plot_type surveyed?

```{r}
surveys %>%
    count(plot_type) 
```

2. Use group_by() and summarize() to find the mean, min, and max hindfoot length for each species (using species_id). Also add the number of observations (hint: see ?n).

```{r}
surveys %>%
    filter(!is.na(hindfoot_length)) %>%
    group_by(species_id) %>%
    summarize(
        mean_hindfoot_length = mean(hindfoot_length),
        min_hindfoot_length = min(hindfoot_length),
        max_hindfoot_length = max(hindfoot_length),
        n = n()
    )
```

3. What was the heaviest animal measured in each year? Return the columns year, genus, species_id, and weight.

```{r}
surveys %>%
    filter(!is.na(weight)) %>%
    group_by(year) %>%
    filter(weight == max(weight)) %>%
    select(year, genus, species, weight) %>%
    arrange(year)
```


## Challenge_4 

1. Spread the surveys data frame with year as columns, plot_id as rows, and the number of genera per plot as the values. You will need to summarize before reshaping, and use the function n_distinct() to get the number of unique genera within a particular chunk of data. It’s a powerful function! See ?n_distinct for more.

```{r}
surveys_spread_genera <- surveys %>%
  group_by(plot_id, year) %>%
  summarize(n_genera = n_distinct(genus)) %>%
  spread(year, n_genera)
```

2. Now take that data frame and gather() it again, so each row is a unique plot_id by year combination.

```{r}
surveys_spread_genera %>%
  gather("year", "n_genera", -plot_id)
```

3. The surveys data set has two measurement columns: hindfoot_length and weight. This makes it difficult to do things like look at the relationship between mean values of each measurement per year in different plot types. Let’s walk through a common solution for this type of problem. First, use gather() to create a dataset where we have a key column called measurement and a value column that takes on the value of either hindfoot_length or weight. 

```{r}
surveys_long <- surveys %>%
  gather("measurement", "value", hindfoot_length, weight)
```

4. With this new data set, calculate the average of each measurement in each year for each different plot_type. Then spread() them into a data set with a column for hindfoot_length and weight. Hint: You only need to specify the key and value columns for spread().

```{r}
surveys_long %>%
  group_by(year, measurement, plot_type) %>%
  summarize(mean_value = mean(value, na.rm=TRUE)) %>%
  spread(measurement, mean_value)
```







