---
title: "Lab 3a"
output: html_document
---

```{r}
library(tidyverse)
```

```{r}
 download.file(url="https://ndownloader.figshare.com/files/2292169",
              destfile = "portal_data_joined.csv")

```
```{r}
surveys <- read.csv("portal_data_joined.csv")

```

## Challenge_1 

Based on the output of str(surveys), can you answer the following questions.
1. What is the class of the object surveys?

Data frame

```{r}
str(surveys)
```

2a. How many columns are in this object?

```{r}
ncol(surveys)
```

2b. How many rows are in this object?

```{r}
nrow(surveys)
```

3. How many species have been recorded during these surveys? 

```{r}
surveys$species_id <- factor(surveys$species_id)
```


```{r}
length(summary(surveys[ , "species_id"]))
```

## Challenge_2

1. Create a data.frame (surveys_200) containing only the data in row 200 of the surveys dataset.

```{r}
surveys_200 <- surveys[200, ]
```

2. Notice how nrow() gave you the number of rows in a data.frame?

2a. Use that number to pull out just that last row in the data frame.

```{r}
n_rows <- nrow(surveys)
```

```{r}
surveys_last <- surveys[n_rows, ]
surveys_last
```

2b. Compare that with what you see as the last row using tail() to make sure it’s meeting expectations.

```{r}
tail(surveys)
```
2c. Pull out that last row using nrow() instead of the row number.

```{r}
nrow(surveys)
```

2d. Create a new data frame (surveys_last) from that last row.

```{r}
surveys_last <- surveys[n_rows, ]
```

3. Use nrow() to extract the row that is in the middle of the data frame. Store the content of this row in an object named surveys_middle.

```{r}
surveys_middle <- surveys[n_rows / 2, ]
surveys_middle
```

4. Combine nrow() with the - notation above to reproduce the behavior of head(surveys), keeping just the first through 6th rows of the surveys dataset.

```{r}
surveys_head <- surveys[-(7:n_rows), ]
```


## Challenge_3

1. Rename “F” and “M” to “female” and “male” respectively.

```{r}
sex <- factor(surveys$sex)
```

```{r}
levels(sex)[2:3] <- c("female", "male")
```

```{r}
sex <- factor(sex, levels = c("female", "male", "undetermined"))
```


2. Now that we have renamed the factor level to “undetermined”, can you recreate the barplot such that “undetermined” is last (after “male”)?

```{r}
plot(sex)
```

```{r}
str(surveys)
```


## Challenge_4

1. We have seen how data frames are created when using read.csv(), but they can also be created by hand with the data.frame() function. There are a few mistakes in this hand-crafted data.frame. Can you spot and fix them? Don’t hesitate to experiment!

animal_data <- data.frame(
          animal = c(dog, cat, sea cucumber, sea urchin),
          feel = c("furry", "squishy", "spiny"),
          weight = c(45, 8 1.1, 0.8)
          )

```{r}
animal_data <- data.frame(
          animal = c("dog", "cat", "sea cucumber", "sea urchin"),
          feel = c("furry", "furry", "squishy", "spiny"),
          weight = c(45, 8, 1.1, 0.8)
)
```

2. Can you predict the class for each of the columns in the following example? Check your guesses using str(country_climate):

2a. Are they what you expected? Why? Why not?

Country, climate, temperature, and northern_hemisphere are factors. has_kangaroo is numeric. 
I expected temperature to be numeric and it is not. 


```{r}
country_climate <- data.frame(
       country = c("Canada", "Panama", "South Africa", "Australia"),
       climate = c("cold", "hot", "temperate", "hot/temperate"),
       temperature = c(10, 30, 18, "15"),
       northern_hemisphere = c(TRUE, TRUE, FALSE, "FALSE"),
       has_kangaroo = c(FALSE, FALSE, FALSE, 1)
       )
```

```{r}
str(country_climate)

```


2b. What would have been different if we had added stringsAsFactors = FALSE when creating the data frame?

stringsAsFactors = FALSE would have made character vectors instead of factors. stringsAsFactors = TRUE would make them factors, shown below. 


```{r}
country_climate <- data.frame(
       country = c("Canada", "Panama", "South Africa", "Australia"),
       climate = c("cold", "hot", "temperate", "hot/temperate"),
       temperature = c(10, 30, 18, "15"),
       northern_hemisphere = c(TRUE, TRUE, FALSE, "FALSE"),
       has_kangaroo = c(FALSE, FALSE, FALSE, 1), stringsAsFactors = TRUE)
```

```{r}
str(country_climate)
```

2c. What would you need to change to ensure that each column had the accurate data type?

```{r}
country_climate <- data.frame(
       country = c("Canada", "Panama", "South Africa", "Australia"),
       climate = c("cold", "hot", "temperate", "hot/temperate"),
       temperature = c(10, 30, 18, 15),
       northern_hemisphere = c(TRUE, TRUE, FALSE, FALSE),
       has_kangaroo = c(FALSE, FALSE, FALSE, 1)
       )
```

```{r}
str(country_climate)
```



